CrypTool Script \\ Mathematics and Cryptography

### 2003-03-11 script/en/chapters/cryptomethods.tex
be updates

% ..............................................................................
%                V E R S C H L U E S S E L U N G S V E R F A H R E N
%
% Schreibregelung: Capitel header to capitalize words,
%                  All other headers in normal lower/upper case 
%                  like sentences without dot at the end.
%
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

The purpose of encryption \index{Encryption} is to change data in such a way
that only an authorised recipient is able to reconstruct the plaintext. This has
the advantage that you can transmit encrypted data openly and nevertheless need
not fear a perpetrator reading the data without authorisation. Authorised
recipients possess a piece of secret information --- called the key --- which
allows them to decrypt the data while it remains hidden from everyone else.\par \vskip + 3pt

One encryption procedure has been proved to be secure --- the {\em One Time
  Pad}.
\index{One ­Time ­Pad} However, this procedure has several practical
disadvantages (the key used must be selected randomly and must be just as long
as the message to be protected), which means that it is hardly used except in
closed environments such as for the hot wire between Moscow and Washington.\par \vskip + 3pt

@@

For {\em symmetric} encryption \index{Encryption!symmetric} the sender and
recipient must possess a common (secret) key which they have exchanged before
actually starting to communicate. The sender uses this key to encrypt the
message and the recipient uses it to decrypt it.\par \vskip + 3pt

@@

Up-to-the-minute procedures are the symmetric AES procedures. The associated
Rijndael procedure was declared winner the AES award on 2 October 2000 and thus
succeeds the DES procedure.

More details about the AES algorithms can be found within the 
Online help of CrypTool\index{CrypTool}
  \footnote{%
      CrypTool Online help\index{CrypTool}: the index see head-word {\bf AES}
      leads to the help pages: {\bf AES candidates}, 
      {\bf The AES Winner Rijndael} and 
      {\bf The Rijndael encryption algorithm}.
  }
.

% --------------------------------------------------------------------------
\subsubsection{New results about cryptanalysis of AES}

@@

AES with a minimum key length of 128 bit is still in the long run sufficiently secure against brute-force attacks - as long as the quantum computers weren't powerful enough. When announced as new standard AES was immune against all known crypto attacks, mostly based on statistical considerations and earlier applied to DES: using pairs of clear and cipher texts expressions are constructed, which are not completely at random, so they allow conclusions to the used keys. These attacks required unrealistically large amounts of intercepted data.

@@

Such a procedure can be demonstrated using a series of thief-proof letter boxes.
If I have composed a message, I then look for the letter box of the recipient
and post the letter through it. After that, I can no longer read or change the
message myself, because only the legitimate recipient possesses the key for the
letter box.\par \vskip + 3pt

The advantage of asymmetric procedures is the easy \index{Key management} key management. Let's look again at a network with $n$
subscribers. In order to ensure that each subscriber can establish
an encrypted connection to each other subscriber, each subscriber
must possess a pair of keys. We therefore need $2n$ keys or $n$
pairs of keys. Furthermore, no secure channel is needed before
messages are transmitted, because all the information required in
order to communicate confidentially can be transmitted openly. In
this case, you simply have to pay attention to the accuracy
(integrity and authenticity) \index{Authenticity} of the public
key. Disadvantage: Pure asymmetric procedures take a lot longer to
perform than symmetric ones.\par \vskip + 3pt

@@

In order to benefit from the advantages of symmetric and asymmetric techniques
together, hybrid procedures are usually used (for encryption) in practice.
\par \vskip + 3pt

In this case the data is encrypted using symmetric procedures: the key is a
session key\index{Session key} generated by the sender
randomly \footnote{%
An important part of cryptographically secure techniques is to generate 
random numbers. Within CrypTool\index{CrypTool} v1.3 you can check out
different random number generators using the menu {\bf Indiv. Procedures 
\textbackslash{} Generate Random Numbers}. 
Using the menu{\bf Analysis \textbackslash{} Random Tests} you can apply
different test methods for random data to binary documents. \\
Till now CrypTool concentrates on cryptographically strong 
pseudo number generators. Only the integrated Secude generator 
involves a "pure" random source. 
}\index{Random}
that is only used for this message.
This session key is then encrypted using the asymmetric procedure and
transmitted to the recipient together with the message. Recipients can determine
the session key using their secret keys and then use the session key to encrypt
the message. In this way, we can benefit from the easy key management
\index{Key management} of asymmetric procedures and encrypt large quantities of data
quickly and efficiently using symmetric procedures.

@@

Beside information you can find in many books and on a lot of websites the 
online help of CrypTool\index{CrypTool} also offers very many details about
the symmetric and asymmetric encryption methods.

>>>>

% ..............................................................................
%                V E R S C H L U E S S E L U N G S V E R F A H R E N
%
% Schreibregelung: Capitel header with all nouns/verbs-words capitalized,
%                  All other headers in the normal lower/upper case manner
%                  like sentences without dot at the end.
%
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

The purpose of encryption \index{Encryption} is to change data in such a way
that only an authorised recipient is able to reconstruct the plaintext. This
has the advantage of being able to transmit encrypted data openly and 
nevertheless need not fear a perpetrator reading the data without 
authorisation. Authorised recipients possess a piece of secret 
information -- called the key -- which allows them to decrypt the data
while it remains hidden from everyone else.\par \vskip + 3pt

One encryption procedure has been proved to be secure -- the {\em One
Time Pad}\index{One ­Time ­Pad}. However, this procedure has several practical
disadvantages (the key used must be selected randomly and must be just as long
as the message to be protected), which means that it is hardly used except in
closed environments such as for the hot wire between Moscow and Washington.\par \vskip + 3pt

@@

For {\em symmetric} encryption \index{Encryption!symmetric} the sender and
recipient must be in possession of a common (secret) key which they have
exchanged before actually starting to communicate. The sender uses this
key to encrypt the message and the recipient uses it to decrypt it.
\par \vskip + 3pt

@@

Up-to-the-minute procedure is the symmetric AES standard. The associated
Rijndael algorithm was declared winner of the AES award on 2 October 2000
and thus succeeds the DES procedure.

More details about the AES algorithms can be found within the 
online help of CrypTool\index{CrypTool}%
\footnote{%
      CrypTool Online help\index{CrypTool}: the index head-word {\bf AES}
      leads to the 3 help pages: {\bf AES candidates}, 
      {\bf The AES winner Rijndael} and 
      {\bf The Rijndael encryption algorithm}.
  }
.

% --------------------------------------------------------------------------
\subsubsection{New results about cryptanalysis of AES}
\index{Cryptanalysis}

@@

AES with a minimum key length of 128 bit is still in the long run sufficiently secure against brute-force attacks -- as long as the quantum computers aren't powerful enough. When announced as new standard AES was immune against all known crypto attacks, mostly based on statistical considerations and earlier applied to DES: using pairs of clear and cipher texts expressions are constructed, which are not completely at random, so they allow conclusions to the used keys. These attacks required unrealistically large amounts of intercepted data.

@@

Such a procedure can be demonstrated using a series of thief-proof letter boxes.
If I have composed a message, I then look for the letter box of the recipient
and post the letter through it. After that, I can no longer read or change the
message myself, because only the legitimate recipient has the key for the
letter box.\par \vskip + 3pt

The advantage of asymmetric procedures is the easy \index{Key management} key management. Let's look again at a network with $n$
subscribers. In order to ensure that each subscriber can establish
an encrypted connection to each other subscriber, each subscriber
must possess a pair of keys. We therefore need $2n$ keys or $n$
pairs of keys. Furthermore, no secure channel is needed before
messages are transmitted, because all the information required in
order to communicate confidentially can be sent openly. In
this case, you simply have to pay attention to the accuracy
(integrity and authenticity) \index{Authenticity} of the public
key. Disadvantage: Pure asymmetric procedures take a lot longer to
perform than symmetric ones.\par \vskip + 3pt

@@

In order to benefit from the advantages of symmetric and asymmetric
techniques together, hybrid procedures \index{Encryption!hybrid} are
usually used (for encryption) in practice. \par \vskip + 3pt

In this case the data is encrypted using symmetric procedures: the key is a
session key\index{Session key} generated by the sender
randomly\footnote{%
An important part of cryptographically secure techniques is to generate 
random numbers. Within CrypTool\index{CrypTool} v1.3 you can check out
different random number generators using the menu {\bf Indiv. Procedures 
\textbackslash{} Generate Random Numbers}. 
Using the menu{\bf Analysis \textbackslash{} Random Tests} you can apply
different test methods for random data to binary documents. \\
Up to now CrypTool has concentrated on cryptographically strong 
pseudo number generators. Only the integrated Secude generator 
involves a "pure" random source. 
}\index{Random}
that is only used for this message.
This session key is then encrypted using the asymmetric procedure and
transmitted to the recipient together with the message. Recipients can determine
the session key using their secret keys and then use the session key to encrypt
the message. In this way, we can benefit from the easy key management
\index{Key management} of asymmetric procedures and encrypt large quantities of data
quickly and efficiently using symmetric procedures.

@@

Beside the information you can find in many books and on a good number 
of websites the online help of CrypTool\index{CrypTool} also offers very
many details about the symmetric and asymmetric encryption methods.

### 2003-03-11 script/en/chapters/introduction.tex
be updates

% --------------------------------------------------------------------------
\section*{Preface to the 6th edition of the CrypTool script}  \addcontentsline{toc}{section}{Preface to the 6th edition of the CrypTool script}

Starting in the year 2000 this script became a part of the 
CrypTool\index{CrypTool} package. It should accomplish the program 
CrypTool by explaining some mathematical topics in more details, 
but still in a good to understand manner.

In order to also enable independent developers/authors work together 
the topics have been split up and for each topic an extra chapter was 
written readable by its own. The later editorial work in TeX added 
cross linkages and descriptions, where you can find the according 
functions within the CrypTool\index{CrypTool} program. Naturally 
there are much more interesting topics in mathematics and cryptography 
which could be more deepened -- therefore this is only one of many
ways to do it.

Rapid spread of the Internet also lead to an intensified research 
in the technologies involved. Especially within the area of cryptography 
much new knowledge came up.

This edition of the script updates the summaries of the following 
topical research areas:         
\begin{itemize}
\item the search for the largest prime numbers (generalized Mersenne 
      and Fermat primes), 
\item progress in number theory ("Primes in P"), 
\item the factorisation of big numbers (TWIRL) and 
\item progress in cryptanalysis of the AES standard. 
% \item usage of elliptic curves.
\end{itemize}

The first time when the script was delivered with CrypTool\index{CrypTool} 
was in version 1.2.01. Since then it was extended and updated with each 
new version of CrypTool (1.2.02, 1.3.00, 1.3.02, 1.3.03 and now 1.3.04).

@@

I am deeply grateful to all the people helping with their impressive engagement to make this project successful and wide-spreaded.

I hope that many readers have fun with this script and that they get 
out of it more interest and higher understanding for this modern but 
also very ancient topic.

@@

% --------------------------------------------------------------------------
\newpage
\section*{Introduction -- How do the script and CrypTool play together?}  \addcontentsline{toc}{section}{Introduction -- How do the script and CrypTool play together?}

@@

The {\em authors} tried to describe cryptography for a broad audience --
without being mathematically incorrect. We believe, that this didactical
pretension is the best way to promote the awareness for IT security and the
readiness to use standardised modern cryptography.
\par \vskip + 15pt

@@

A further aim was to enable users to understand the cryptographic procedures
implemented in the Deutsche Bank. In this way, using CrypTool as a reliable
reference implementation of the various encryption procedures (because of using the industry-proven Secude Library\index{Secude}),
you can test the encryption implemented in other programs. \par \vskip + 3pt

Since then CrypTool\index{CrypTool} has often been used for 
education in companies and teaching and several universities 
help to further develop the project. \par \vskip + 3pt
\vskip +10 pt

At this point I'd like to thank explicitely 3 persons who especially 
contributed to CrypTool\index{CrypTool}. Without their talents and engagement 
CrypTool would not be what it is today:
\begin{itemize}
   \item Mr. Henrik Koy
   \item Mr. J"org-Cornelius Schneider and
   \item Dr. Peer Wichmann.
\end{itemize}
Also I want to thank all the many ones not named here for their 
engagement (mostly performed in their spare time).
\\

>>>>

% --------------------------------------------------------------------------
\section*{Preface to the 6th Edition of the CrypTool Script}  \addcontentsline{toc}{section}{Preface to the 6th Edition of the CrypTool Script}

Starting in the year 2000 this script became a part of the 
CrypTool\index{CrypTool} package. It should accomplish the program 
CrypTool by explaining some mathematical topics in more details, 
but still in a way which is easy to.

In order to also enable independent developers/authors work together 
the topics have been split up and for each topic an extra chapter was 
written readable by its own. The later editorial work in TeX added 
cross linkages between different sections and footnotes, where you

can find the according 
functions within the CrypTool\index{CrypTool} program. Naturally 
there are much more interesting topics in mathematics and cryptography 
which could be more expanded -- therefore this is only one of many
ways to do it.

Rapid spread of the Internet has also lead to an intensified research 
in the technologies involved, especially within the area of cryptography 
where a good deal of new knowledge came up.

This edition of the script updates the summaries of the following 
topical research areas:
\vspace{-7pt}
\begin{itemize}
  \item the search for the largest prime numbers (generalized Mersenne 
        and Fermat primes), 
  \item progress in number theory ("Primes in P"), 
  \item the factorisation of big numbers (TWIRL) and 
  \item progress in cryptanalysis of the AES standard. 
% \item usage of elliptic curves [not only over GP(p)].
\end{itemize}

The first time the script was delivered with CrypTool\index{CrypTool} 
was in version 1.2.01. Since then it has been extended and updated with each 
new version of CrypTool (1.2.02, 1.3.00, 1.3.02, 1.3.03 and now 1.3.04).

@@

I am deeply grateful to all the people helping with their impressive
committment to make this project successful and wide-spreading.

I hope that many readers have fun with this script and that they get 
out of it more interest and greater understanding for this modern but 
also very ancient topic.
\\

@@

% --------------------------------------------------------------------------
\newpage
\section*{Introduction -- How do the Script and CrypTool Play together?}  \addcontentsline{toc}{section}{Introduction -- How do the Script and CrypTool Play together?}

@@

The {\em authors} have attempted to describe cryptography for a broad 
audience -- without being mathematically incorrect. We believe that this
didactical pretension is the best way to promote the awareness for IT
security and the readiness to use standardised modern cryptography.
\par \vskip + 15pt

@@

A further aim has been to enable users to understand the cryptographic
procedures implemented in the Deutsche Bank. In this way, using CrypTool
as a reliable reference implementation of the various encryption procedures
(because of using the industry-proven Secude Library\index{Secude}),
you can test the encryption implemented in other programs. \par \vskip + 3pt

Since then CrypTool\index{CrypTool} has often been used for 
education in companies and teaching at school and universities and
moreover several universities are helping to further develop the project.
\par \vskip + 15pt

\textbf{Acknowledgment}

At this point I'd like to thank explicitely 3 people who especially 
contributed to CrypTool\index{CrypTool}. Without their talents and engagement 
CrypTool would not be what it is today:
\vspace{-7pt}
\begin{itemize}
   \item Mr. Henrik Koy
   \item Mr. J\"org-Cornelius Schneider and
   \item Dr. Peer Wichmann.
\end{itemize}
Also I want to thank all the many people not mentioned here for their 
hard work (mostly carried out in their spare time).
\\

### 2003-03-11 script/en/chapters/numbertheory.tex
be updates

This ``introduction'' is for people mathematically interested. There is no more pre-knowledge necessary than
what you learn in the secondary school.\par

We by will had ``beginners'' in mind; we did not take the approach of mathematical textbooks, called 
``introduction'', which can not be understood at the first go further than page 5 and
which have the real purpose to deliver all information that special monographs can be read.

@@

The goal is not to exhaustively explain the number theory findings, but to show the essential
procedures. The volume of the content is oriented that the reader can understand and apply the RSA method.

@@

Number theory has always been a source of fascination -- for both amateurs and professional mathematicians. In 
contrast to other areas of mathematics, many of the problems and theorems in number theory can be understood 
by non-experts. On the other hand, mathematicians often take a long time to find solutions to the problems or 
prove the theorems. It is therefore easy to pose good questions but quite another matter to find the answer. One 
example of this is what is known as Fermat’s Last (or large) theorem\footnote{One of the things you learn in mathematics at school is Pythagoras theorem, which states the following for a right-angle 
triangle: $a^2 + b^2 = c^2$, where $a$ and $b$ are the lengths of the sides containing the right angle and $c$ is the 
length of the hypotenuse. Fermat famously proposed that $a^n + b^n \not= c^n$ for whole-number exponents $n 
> 2$. Unfortunately, the letter in which Fermat made the claim did not have enough space for him to prove it. 
The theorem was not proved until over 300 years later \cite[p. 433-551]{Wiles1994}\index{Wiles}. }.

@@

% --------------------------------------------------------------------------
\newpage
\subsubsection{Convention}
Unless stated otherwise: 
\begin{itemize}
\item The letter $a, b, c, d, e, k, n, m, p, q$ are used to present whole numbers.
\item The letter $i$ and $j$ represent natural numbers. 
\item The letter $p$ always represents a prime number.
\item Te sets $\mathbb{N} = \{ 1, 2, 3, \cdots \}$ and $\mathbb{Z} =\{ \cdots, -3, -2, -1, 0, 1, 2, 3, \cdots \}$ 
are the {\em natral} and {\em whole} numbers respectively.
\end{itemize}

@@

\item[{\bf 1.}] Select two distinct random prime numbers
\footnote{Compaq introduced the so-called multi-prime method with high marketing 

@@

\href{http://www.bsi.de/aufgaben/projekte/pbdigsig/download/kryptalg.pdf}{\tt http://www.bsi.de/aufgaben/projekte/pbdigsig/download/kryptalg.pdf}).}.

@@

                $\max(p,q)  <  e <  J(n) - 1$
                \footnote{The procedure also allows us to select $d$ freely and then 

@@

\item[{\bf 3.}] Select $d \in \{1, \cdots, n-1\}$ with  
                $e*d \equiv 1  {\rm ~(mod~} J(n))$,
                i.e. $d$ is the multiplicative inverse of $e$ modulo $J(n)$
                \footnote{For reasons of security, $d$ should not 
                be too small.}$^,$
                \footnote{We start by determining either $d$ or $e$ depending 

@@

Alice and Bob\footnote{Bob and Alice are the standard names used for the two authorised participants in a protocol 
(see \cite[p. 23]{Schneier1996}).
} use a one way function to obtain a key $S$, the session key, for subsequent correspondence. This is then a secret 
that is only known to the two of them. Alice selects a random number $a$ and keeps it secret. She applies a one way 
function to $a$ to calculate the number $A = g^a$ and sends it to Bob. He does the same, by selecting a secret random 
number $b$, calculating $B = g^b$ and sending it to Alice. The number $g$ is random and can be publicly known. 
Alice applies the one way function together with her secret number $a$ to $B$, while Bob does the same with his 
secret number $b$ and the received number $A$.

@@

Before applying the RSA procedure to a text, we will first demonstrate it directly
using a number
\footnote{Using CrypTool\index{CrypTool} v1.3 you can solve this with
the {\bf Indiv.Procedures \textbackslash{} RSA Demonstration \textbackslash{} RSA Cryptosystem}.}.
\begin{itemize}
\item[{\bf 1.}] Let the selected prime numbers be $p=5$ and $q=11$. \\
Thus, $n=55$ and $J(n)=(p-1)*(q-1)=40$.
\item[{\bf 2.}] $e = 7$ (should lie between $11$ and $40$ and must be relatively prime to $40$).
\item[{\bf 3.}] $d = 23$ (since $23*7 \equiv 161 \equiv 1 {\rm ~(mod~} 40)$),
    \begin{itemize}
    \item[] $\rightarrow$ Public key of the recipient:  $(55, 7),$
    \item[] $\rightarrow$ Private key of the recipient: $(55, 23).$
    \end{itemize}
\item[{\bf 4.}] Let the message be the number $M = 2$ (so no division into blocks is required).
\item[{\bf 5.}] Encryption: $C \equiv 2^7 \equiv 18 {\rm ~(mod~}55).$
\item[{\bf 6.}] The cipher text is simply the number $C = 18$ (we therefore do not need to divide it into blocks).
\item[{\bf 7.}] Decryption: 
        $M \equiv 18^{23} \equiv 18^{(1+2+4+16)} \equiv 18*49*36*26 \equiv 2 {\rm ~(mod~}55).$
\end{itemize}

@@

Using the selected values it is easy for a cryptanalyst to derive the secret values from the public 
parameters $n=3,713$ and $e=37$ by revealing that $3,713 = 47 * 79$.\\  
If $n$ is a $768$-bit number, there is, according to present knowledge, little chance of this.

@@

However, it is not the 
result that is important here but rather the individual steps of the solution, that is, the explanation of the 
cryptanalysis\footnote{The method of solving the problem is outlined in the scenario of the online help to CrypTool\index{CrypTool} 
and in the presentation on the website. If anyone sends us a well prepared exact method of solving the problem, we 
would be pleased to include it in the documentation.}.

@@

The following task is a corrected version from the excellent book written by Prof. Yan 
\cite[Example 3.3.7, p. 318]{Yan2000}\index{Yan 2000}.
However, it is not the result that is important here but rather the individual steps of the solution, that is, 
the explanation of the cryptanalysis\footnote{The method of solving the problem is outlined in the scenario of 
the online help to CrypTool\index{CrypTool} and in the CrypTool presentation. If anyone sends us a well prepared exact 
method of solving the problem, we would be pleased to include it in the documentation.}.

@@

\bibitem[GISA2001xxx?]{GISA2001xxx?} \index{GISA}
    GISA (German Information Security Agency) \\
    {\em Recommendation for key length selection},\\
    \href{http://www.bsi.de/aufgaben/projekte/pbdigsig/download/kryptalg.pdf}
    {\tt http://www.bsi.de/aufgaben/projekte/pbdigsig/download/kryptalg.pdf} \\
    A statement on these recommendations: \\
    \href{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}
    {\texttt{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}}

@@

   \item Das Cunningham-Projekt, \index{Cunningham-Projekt}\\ 
         \href{http://www.cerias.purdue.edu/homes/ssw/cun/}
         {\texttt{http://www.cerias.purdue.edu/homes/ssw/cun/}}

@@

I would like to take this opportunity to thank 
\begin{itemize}
  \item {Mr. Henrik Koy for making many very useful suggestions, 
         for very constructively proof-reading this article and
         for helping with TeX.}
  \item {J\"org-Cornelius Schneider for his TeX support.}
  \item {Dr. Illies for pointing me to Pari-GP \index{Pari-GP}.}
\end{itemize}

>>>>

This ``introduction'' is for people with a mathematical interest. There is no
more pre-knowledge necessary than what you learn in the secondary school.\par

We intentionally had ``beginners'' in mind; we did not take the approach
of mathematical textbooks, called ``introduction'', which cannot be
understood at the first reading further than page 5 and which have the real
purpose to deliver all information that special monographs can be read.

@@

The goal is not to exhaustively explain the number theory findings, but 
to show the essential procedures. The volume of the content is so oriented
that the reader can understand and apply the RSA method.

@@

Number theory has always been a source of fascination -- for both amateurs and professional mathematicians. In 
contrast to other areas of mathematics, many of the problems and theorems in number theory can be understood 
by non-experts. On the other hand, mathematicians often take a long time to find solutions to the problems or 
prove the theorems. It is therefore one thing to pose good questions but quite another matter to find the answer. One 
example of this is what is known as Fermat's Last (or large) theorem\footnote{One of the things you learn in mathematics at school is Pythagoras theorem, which states the following for a right-angle 
triangle: $a^2 + b^2 = c^2$, where $a$ and $b$ are the lengths of the sides containing the right angle and $c$ is the 
length of the hypotenuse. Fermat famously proposed that $a^n + b^n \not= c^n$ for whole-number exponents $n 
> 2$. Unfortunately, the letter in which Fermat made the claim did not have enough space for him to prove it. 
The theorem was not proved until over 300 years later \cite[p. 433-551]{Wiles1994}\index{Wiles}. }.

@@

% --------------------------------------------------------------------------
\newpage
\subsubsection{Convention}
Unless stated otherwise: 
\begin{itemize}
\item The letter $a, b, c, d, e, k, n, m, p, q$ are used to present whole numbers.
\item The letter $i$ and $j$ represent natural numbers. 
\item The letter $p$ always represents a prime number.
\item Te sets $\mathbb{N} = \{ 1, 2, 3, \cdots \}$ and $\mathbb{Z} =\{ \cdots, -3, -2, -1, 0, 1, 2, 3, \cdots \}$ 
are the {\em natural} and {\em whole} numbers respectively.
\end{itemize}

@@

\item[{\bf 1.}] Select two distinct random prime numbers\footnote{
Compaq introduced the so-called multi-prime method with high marketing 

@@

\href{http://www.bsi.bund.de/esig/basics/techbas/krypto/bund02v7.pdf}
{\tt http://www.bsi.bund.de/esig/basics/techbas/krypto/bund02v7.pdf}).
}.

@@

                $\max(p,q)  <  e <  J(n) - 1$\footnote{The 
		procedure also allows us to select $d$ freely and then 

@@

\item[{\bf 3.}] Select $d \in \{1, \cdots, n-1\}$ with $e*d \equiv 1  
                {\rm ~(mod~} J(n))$, i.e. $d$ is the multiplicative inverse
		of $e$ modulo $J(n)$\footnote{For reasons of security, 
		$d$ should not be too small.}$^,$\footnote{We start by determining either $d$ or $e$ depending 

@@

Alice and Bob\footnote{Bob and Alice are the default names used for the two authorised participants in a protocol 
(see \cite[p. 23]{Schneier1996}).
} use a one way function to obtain a key $S$, the session key, for subsequent correspondence. This is then a secret 
that is only known to the two of them. Alice selects a random number $a$ and keeps it secret. She applies a one way 
function to $a$ to calculate the number $A = g^a$ and sends it to Bob. He does the same, by selecting a secret random 
number $b$, calculating $B = g^b$ and sending it to Alice. The number $g$ is random and can be publicly known. 
Alice applies the one way function together with her secret number $a$ to $B$, while Bob does the same with his 
secret number $b$ and the received number $A$.

@@

Before applying the RSA procedure to a text, we will first demonstrate it
directly using a single number as message
\footnote{Using CrypTool\index{CrypTool} v1.3 you can solve this with
the {\bf Indiv.Procedures \textbackslash{} RSA Demonstration \textbackslash{} RSA Cryptosystem}.}.
\begin{itemize}
\item[{\bf 1.}] Let the selected prime numbers be $p=5$ and $q=11$. \\
Thus, $n=55$ and $J(n)=(p-1)*(q-1)=40$.
\item[{\bf 2.}] $e = 7$ (should lie between $11$ and $40$ and must be relatively prime to $40$).
\item[{\bf 3.}] $d = 23$ (since $23*7 \equiv 161 \equiv 1 {\rm ~(mod~} 40)$),
    \begin{itemize}
    \item[] $\rightarrow$ Public key of the recipient:  $(55, 7),$
    \item[] $\rightarrow$ Private key of the recipient: $(55, 23).$
    \end{itemize}
\item[{\bf 4.}] Let the message be the number $M = 2$ (so no division into blocks is required).
\item[{\bf 5.}] Encryption: $C \equiv 2^7 \equiv 18 {\rm ~(mod~}55).$
\item[{\bf 6.}] The cipher text is simply the number $C = 18$ (we therefore do not need to divide it into blocks).
\item[{\bf 7.}] Decryption: 
        $M \equiv 18^{23} \equiv 18^{(1+2+4+16)} \equiv 18*49*36*26 \equiv 2 {\rm ~(mod~}55).$
\end{itemize}

@@

Using the selected values it is easy for a cryptanalyst\index{Cryptanalysis}
to derive the secret values from the public 
parameters $n=3,713$ and $e=37$ by revealing that $3,713 = 47 * 79$.\\  
If $n$ is a $768$-bit number, there is, according to present knowledge, little chance of this.

@@

However, it is not the 
result that is important here but rather the individual steps of the solution, that is, the explanation of the 
cryptanalysis\index{Cryptanalysis}\footnote{The method of solving the problem is outlined in the scenario of the online help to CrypTool\index{CrypTool} 
and in the presentation on the website. If anyone sends us a well prepared exact method of solving the problem, we 
would be pleased to include it in the documentation.}.

@@

The following task is a corrected version from the excellent book written by Prof. Yan 
\cite[Example 3.3.7, p. 318]{Yan2000}\index{Yan 2000}.
However, it is not the result that is important here but rather the individual steps of the solution, that is, 
the explanation of the cryptanalysis\index{Cryptanalysis}\footnote{The method of solving the problem is outlined in the scenario of 
the online help to CrypTool\index{CrypTool} and in the CrypTool presentation. If anyone sends us a well prepared exact 
method of solving the problem, we would be pleased to include it in the documentation.}.

@@

\bibitem[GISA2002]{GISA2002} \index{GISA 2002}
    GISA (German Information Security Agency) \\
    {\em Recommendation for key length selection}, Bonn, Sep. 2002, \\
    \href{http://www.bsi.de/aufgaben/projekte/pbdigsig/download/kryptalg.pdf}
    {\tt http://www.bsi.de/aufgaben/projekte/pbdigsig/download/kryptalg.pdf} \\
    A statement on these recommendations: \\
    \href{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}
    {\texttt{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}}

@@

   \item The Cunningham Project, \index{Cunningham project}\\ 
         \href{http://www.cerias.purdue.edu/homes/ssw/cun/}
         {\texttt{http://www.cerias.purdue.edu/homes/ssw/cun/}}

@@

I would like to take this opportunity to thank 
\begin{itemize}
  \item {Mr. Henrik Koy for making many very useful suggestions, 
         for very constructive proof-reading this article and
         for helping with TeX.}
  \item {J\"org-Cornelius Schneider for his TeX support.}
  \item {Dr. Illies for pointing me to Pari-GP \index{Pari-GP}.}
\end{itemize}
